Interest Flooding Attack and Countermeasures in Named Data Networking

### 2013-01-15 03-interest-flooding.tex
Small reorganization of evaluation section. Some spacing changes
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@915 df1a90cd-247b-4364-bdaf-287a704eb499
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@930 df1a90cd-247b-4364-bdaf-287a704eb499

Unlike Data packets, Interest packets in NDN are unsolicited, consume state in intermediate routers, and are routed through the network based on their content name prefix. These properties make Interest packets an effective tool in launching DoS and DDoS attacks in NDN. An attacker or a set of distributed attackers can inject excessive amounts of Interests to overload the network and cause service disruptions for legitimate users (Fig.~\ref{fig:flooding example}). 
%%% Alex %%% We have (almost) exactly the same sentence in the intro. I think here is kind of redundant, as it is obvious how we call the attack from the section title.  Also, we are referring to the term again in the last paragraph of the section.
% We coin the term \emph{Interest flooding} to refer to these network-level flooding attacks.

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.39]{attack-definition}
  \caption{Example of the Interest flooding attack}
  \label{fig:flooding example}
\end{figure}

@@

Since communication in NDN is content-centric, it is unlikely that an adversary can target specific routers or end-hosts, as neither is required to have a globally routable network identity/address. However, an adversary can target a specific namespace.
For example in Fig.~\ref{fig:flooding example}, if the Data producer is the exclusive owner of \ndnName{/foo/bar} namespace, both router B and the Data producer would receive all Interests for \ndnName{/foo/bar/\ldots} that cannot be otherwise satisfied from in-network caches.%
\footnote {In our example, Data producer is single-homed. In the case of multi-homed producers with rich connectivity through multiple providers, NDN would likely to cope better with DDoS attacks due its native multipath and adaptive forwarding~\cite{adaptive-forwarding}.}
A large volume of these Interests can distrupt service quality in NDN network in two ways: \emph{create network congestion} and \emph{exhaust resources on routers}.

Similar to packets in traditional networks, Interest packets in NDN consume a portion of network capacity. A large number of Interest packets might cause congestion and lead to legitimate packets being dropped in the network. Although congestion might occur anywhere in the network, a coordinated DDoS attack would likely target one specific namespace to concentrate attack traffic in certain segments of the network, typically closest to the publisher serving that namespace since name prefixes in NDN loosely correspond to a network location or sets of network locations.

As NDN routers maintain per-packet states for each forwarded Interest (i.e., an entry in its PIT), an excessive amount of malicious Interests can lead to exhaustion of a routers' memory. When a router has insufficient memory to create a new PIT entry, it is forced to drop new incoming Interests, thus disrupting service for legitimate users.
% there is no point in forwarding the Interest further.
% Otherwise, if Data comes back, it would be considered unsolicited and dropped anyways.
%Therefore, service for legitimate users can be disrupted due to limited memory resources in routers without saturating communication links to their limits.

@@

To efficiently implement an Interest flooding attack targeting a specific namespace (e.g., \ndnName{/newyorktimes/}), an adversary needs to make sure that (1) the expressed Interests are routed as close to the Data producer as possible, and (2) the corresponding PIT entries are stored at intermediate NDN routers for as long as possible.
In order to achieve both requirements effectively, malicious Interests should avoid Interest collapsing and avoid being served from caches of intermediate routers: each malicious Interest generated by an individual attacker should request an unpopular or non-existing content, e.g., content with a unique name (\emph{unique junk Interests}). In this paper, we exclusively focus on this particular attack strategy as it maximizes the damage from each malicious Interest in the network. Other less effective strains of this attack can be mitigated by applying the same or similar countermeasures.

>>>>

As we explained earlier, Interest packets in NDN are routed through the network based on their content name prefixes and consume memory resources at intermediate routers. Thus they can be used to launch DDoS attacks in an NDN network. An attacker or a set of distributed attackers can inject excessive amounts of Interests in an attempt to overload the network and cause service disruptions for legitimate users (Fig.~\ref{fig:flooding example}). 
%%% Alex %%% We have (almost) exactly the same sentence in the intro. I think here is kind of redundant, as it is obvious how we call the attack from the section title.  Also, we are referring to the term again in the last paragraph of the section.
% We coin the term \emph{Interest flooding} to refer to these network-level flooding attacks.

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.39]{attack-definition}
  \vspace{-0.4cm}
  \caption{Example of the Interest flooding attack}
  \label{fig:flooding example}
\end{figure}

@@

Since an NDN network fetches name data, an adversary cannot target specific routers or end-hosts.
%% as neither is required to have a globally routable network identity/address.
However, an adversary can target a specific namespace.
For example in Fig.~\ref{fig:flooding example}, if the Data producer is the exclusive owner of \ndnName{/foo/bar} namespace, both router B and the Data producer would receive all Interests for \ndnName{/foo/bar/\ldots} that cannot be otherwise satisfied from in-network caches.%
\footnote {This example assumes that the adversary floods the network with unique data names carrying \ndnName{/foo/bar} prefix to make them effective. It also assumes the producer is single-homed and the data is not replicated elsewhere. With multi-homed producers or replicated data, NDN would likely to cope better with DDoS attacks due its native multipath and adaptive forwarding~\cite{adaptive-forwarding}.}
A large volume of such malicious Interests can disrupt service quality in NDN network in two ways: \emph{create network congestion} and \emph{exhaust resources on routers}.

Similar to packets in traditional networks, Interest packets in NDN consume a portion of network capacity. A large number of Interest packets might cause congestion and lead to legitimate packets being dropped in the network. In particular a coordinated DDoS attack could target one specific namespace to concentrate attack traffic in certain segments of the network, typically closest to the publisher serving that namespace.
%% since name prefixes in NDN loosely correspond to a network location or sets of network locations.

As NDN routers maintain per-packet states for each forwarded Interest (i.e., an entry in its PIT), an excessive amount of malicious Interests can lead to exhaustion of a router's memory, making the router unable to create new PIT entries for incoming Interests, thus disrupting service for legitimate users.
% there is no point in forwarding the Interest further.
% Otherwise, if Data comes back, it would be considered unsolicited and dropped anyways.
%Therefore, service for legitimate users can be disrupted due to limited memory resources in routers without saturating communication links to their limits.

@@

An effective Interest flooding attack is non-trivial to create.
To target a specific namespace (e.g., \ndnName{/newyorktimes/}), an adversary needs to make sure that (1) the expressed Interests are routed as close to the Data producer as possible, and (2) the corresponding PIT entries are stored at intermediate NDN routers for as long as possible.
%In order to achieve both requirements effectively, malicious Interests should avoid Interest collapsing and 
To avoid an Interest being served from caches of intermediate routers, every single malicious Interest has to request an unpopular or non-existing content, e.g., content with a unique name (\emph{unique junk Interests}). In this paper, we exclusively focus on this particular attack strategy as it maximizes the damage from each malicious Interest in the network. Other less effective strains of Interest flooding attacks can be mitigated by applying the same or similar countermeasures.
